---
title: "Data 612 - Project 1"
author: "Joseph Simone"
date: "2/11/2020"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<STYLE>
table {
    border: 1px solid black;
}
th {
    background-color: rgb(12, 99, 204);
    color: white;
    font-weight: bold;
    padding: 20px 30px;
}
tr:nth-child(even) {
    background-color: rgb(220,220,220);
}
tr:nth-child(odd) {
    background-color: rgb(184, 174, 174);
}
</STYLE>


```{r, message=FALSE, warning= FALSE}
library(dplyr)
library(tidyr)
library(caTools)  
```


### Background

For this project, the recommender system I would like to put into effevt would be one that would recommend Films, more specifically the Star Wars Skywalker Film Franchise. 

This is an industry, where the Ratings or Cristics Reviews for a specific Film could make or break a companies fiscal year. Therefore, there is an abundancy of data to be collected on these various Films. 

### Data-Set

First, I wanted to do was created my own smaller Data-Set of the Star Wars Films using a simple scaling metric to interpert and the same Critics spanning from 1977-2019. Therefore, I decided to use the website [Meta Critic](http://www.metacritic.com). 

Since this website includes various Critics that use the same numerical ratings system of a 0-100 scale. In addition, this gave me the opportunity to scrap real Reviews from the time of the Film's release, instead of using arbitrary values.

### Data Import

```{r section1}
star_wars_data <- read.csv("https://raw.githubusercontent.com/josephsimone/Data-612/master/project_1/star_wars_critic_ratings.csv")  
colnames(star_wars_data) <- gsub("ï..Critics", "Critics", colnames(star_wars_data))
star_wars_data
```


### Splitting Data into Training/Testing Sets 

In this seection, the Data-Frame was first converted to long format. 

Then split into training and testing sets based on 0.75 split ratio.

```{r section2}
split_star_wars_data <- star_wars_data %>% gather(key = Film, value = Review, -Critics)
```


```{r section3}
set.seed(50)
split <- sample.split(split_star_wars_data$Review, SplitRatio = 0.75)
```


```{r section4}
train_data_set <- split_star_wars_data
train_data_set$Review[!split] <- NA
head(train_data_set)
```


```{r section5}
test_data_set <- split_star_wars_data
test_data_set$Review[split] <- NA

head(test_data_set)
```

Since, there are now two different dataset randomly selected, it is time to move onto the RMSE calculations.


```{r section6}
raw_avg <- sum(train_data_set$Review, na.rm = TRUE) / length(which(!is.na(train_data_set$Review)))

rmse_raw_train <- sqrt(sum((train_data_set$Review[!is.na(train_data_set$Review)] - raw_avg)^2) /
                         length(which(!is.na(train_data_set$Review))))
rmse_raw_train
```


```{r section7}
rmse_raw_test <- sqrt(sum((test_data_set$Review[!is.na(test_data_set$Review)] - raw_avg)^2) /
                        length(which(!is.na(test_data_set$Review))))
rmse_raw_test
```

We can observe that RMSE values are significantly larger than expected in a smaller sample space.

### Baseline Predictors

```{r section8}
Critics_bias <- train_data_set %>% filter(!is.na(Review)) %>% 
  group_by(Critics) %>%
  summarise(sum = sum(Review), count = n()) %>% 
  mutate(bias = sum/count-raw_avg) %>%
  select(Critics, CriticsBias = bias)
CriticsBias<-Critics_bias$CriticsBias
```


```{r section9}
Film_bias <- train_data_set %>% filter(!is.na(Review)) %>% 
  group_by(Film) %>%
  summarise(sum = sum(Review), count = n()) %>% 
  mutate(bias = sum/count-raw_avg) %>%
  select(Film, FilmBias = bias)
FilmBias<-Film_bias$FilmBias
```


```{r section10}
train_data_set <- train_data_set %>% left_join(Critics_bias, by = "Critics") %>%
  left_join(Film_bias, by = "Film") %>%
  mutate(RawAvg = raw_avg) %>%
  mutate(Baseline = RawAvg + CriticsBias + FilmBias)
train_data_set
```


```{r section11}
test_data_set <- test_data_set %>% left_join(Critics_bias, by = "Critics") %>%
  left_join(Film_bias, by = "Film") %>%
  mutate(RawAvg = raw_avg) %>%
  mutate(Baseline = RawAvg + CriticsBias + FilmBias)
test_data_set
```


```{r section12}
rmse_base_train <- sqrt(sum((train_data_set$Review[!is.na(train_data_set$Review)] - 
                               train_data_set$Baseline[!is.na(train_data_set$Review)])^2) /
                          length(which(!is.na(train_data_set$Review))))
rmse_base_test <- sqrt(sum((test_data_set$Review[!is.na(test_data_set$Review)] - 
                              test_data_set$Baseline[!is.na(test_data_set$Review)])^2) /
                         length(which(!is.na(test_data_set$Review))))
```

### RMSE

The Largest Bias are determined by the NA Critic's Reviews that did not review the Film.


```{r section13}
rmse_base_train
rmse_base_test
```


The table below represents the RMSE values for both the Training and Testing Sets and the Raw Average and Baseline Predictors.


```{r echo = FALSE}
rmse <- as.data.frame(c(rmse_raw_train, rmse_base_train, rmse_raw_test, rmse_base_test))
colnames(rmse) <- "RMSE"
rownames(rmse) <- c("Training: Raw Average",
                    "Training: Baseline Predictor",
                    "Testing: Raw Average",
                    "Testing: Baseline Predictor")
knitr::kable(rmse) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), 
                            full_width = FALSE)
```

### Summary

The RMSE values improved  the Baseline Predictors in both the Training and Test Sets. 

Even with a smaller Data-Set that included incomplete values, this was enough information to be able to visualize and apply specific Films and Critic bias into our model.
